---
title: "Collect predictions and identify best config"
author: "John Curtin & Claire Punturieri"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  study: "messages"
  batch: "liwc"
  version: "v1"
  cv: "kfold_6_x_5"
  algorithms: "xgboost"   # "all" or name of specific algorithm
editor_options: 
  chunk_output_type: console
---

# Housekeeping

## Code Status



# Set Up

## Set Up Environment

### Study parameters
```{r}
study <- params$study
batch <- params$batch
version <- params$version
cv <- params$cv
algorithms <- params$algorithms
```

### Defaults
```{r}
#| message: false
#| warning: false

options(conflicts.policy = "depends.ok")

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
source("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")

knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')

options(tibble.width = Inf)
options(tibble.print_max = Inf)
```

### Packages and source
```{r}
library(tidyverse)

theme_set(theme_classic())

```

### Paths
```{r}
#| message: false

path_models <- format_path(str_c("risk/models/", study))
path_output <- format_path(str_c("risk/chtc/", study))

```

# Data checking

## Basic checks

Get batch name from above-specified model parameters.
```{r}
batch_name <- str_c("train_", algorithms, "_", cv, "_", version, "_", batch)
```

Load in results file.
```{r}
results <- read_csv(here::here(path_output, batch_name, "output", "batch_results.csv"),
                        show_col_types = FALSE)

```

Conduct final checks -- we expect to see results split evenly among possible hyperparameter combinations, resampling techniques, etc. If these numbers are not even, that suggests there may be a problem.
```{r}
  results |> tab(outer_split_num) |> print()
  results |> tab(inner_split_num) |> print()
  results |> tab(algorithm) |> print()
  results |> tab(feature_set) |> print()
  results |> tab(hp1) |> print()
  results |> tab(hp2) |> print()
  results |> tab(hp3) |> print()
  results |> tab(resample) |> print()
```

## Identify best configuration

Summarize median auROCs by model parameters.
```{r}
metrics_avg <- results |> 
  group_by(algorithm, feature_set, hp1, hp2, hp3, resample) |> 
   summarize(median_roc_auc = median(roc_auc),
             n_folds = n(), .groups = "drop") |> 
  relocate(n_folds) |> 
  arrange(desc(median_roc_auc)) |> 
  ungroup()
```

Check to make sure summarized tibble looks accurate.
```{r}
metrics_avg |> 
  slice(1:50) |> view()
  print_kbl()
```

Pull out best configuration from metrics_avg and save out results for best configuration into models folder on server.
COCO - change the file name, write out csv
```{r}
best_config <- metrics_avg |> 
  slice(1) |> 
  print()

# results |> 
#   filter(algorithm == best_config$algorithm,
#          feature_set == best_config$feature_set,
#          hp1 == best_config$hp1,
#          hp2 == best_config$hp2,
#          hp3 == best_config$hp3,
#          resample == best_config$resample) |> 
#   write_csv(here::here(path_models, str_c("best_config_",
#                                              version, "_",
#                                              cv, ".csv")))
```


## Plot hyperparameters

Ensure that you have assessed a wide enough range of hyperparameters.
```
# update algorithms to actual ones in the tibble
algorithms <- unique(metrics_avg$algorithm) 
feature_sets <- unique(metrics_avg$feature_set) 

for (k in algorithms) {
  
  results_k <- metrics_avg |> 
      filter(algorithm == k)
  
  for (i in feature_sets) {
  
    results_i <- results_k |> 
      filter(feature_set == i)
    
    
    # glmnet
    if (k == "glmnet") {
  
      plot_title <- str_c("Plotting glmnet hyperparameters for ", i, " feature set")
  
  
      plot_i <- results_i |>
        mutate(hp1 = factor(hp1, ordered = TRUE),
               resample = case_when(resample == "none" ~ "none_none",
                                    TRUE ~ resample)) |> 
        separate(resample, c("resample", "under_ratio"), "_") |> 
        mutate(under_ratio = factor(under_ratio)) |> 
        ggplot(mapping = aes(x = log(hp2), 
                         y = roc_auc, 
                         group = hp1, 
                         color = hp1)) +
          geom_line() +
          facet_grid(under_ratio ~ resample) +
          scale_color_discrete(name = "mixture (alpha)") +
          labs(title = plot_title, x = "penalty (lambda)", y = "ROC AUC")
  
      print(plot_i)
    }


    # random forest
    if (k == "random_forest") {
      
      plot_title <- str_c("Plotting RF hyperparameters for ", i, " feature set")
      
      plot_i <- results_i |>
        mutate(hp2 = factor(hp2, ordered = TRUE)) |> 
        separate(resample, c("resample", "under_ratio"), "_") |> 
        mutate(under_ratio = factor(under_ratio)) |> 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(under_ratio ~ resample) +
          scale_color_discrete(name = "min n") +
          labs(title = plot_title, x = "mtry", y = "ROC AUC")
      
       print(plot_i)
    }  
    
    # XGBoost
    if (k == "xgboost") {
      
      # hp1 x axis
            plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and DOWNSAMPLE")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)), # prob have to change this to factor
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "down")) |> 
        ggplot(mapping = aes(x = log_hp1, 
                         y = median_roc_auc, 
                         group = hp3, 
                         color = hp3)) +
          geom_line() +
          facet_grid(resample ~ hp2) +
          scale_color_discrete(name = "mtry") +
          labs(title = plot_title, x = "log10 learning rate", y = "ROC AUC")
      
       print(plot_i)
       
      plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and UPSAMPLE")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "up")) |> 
        ggplot(mapping = aes(x = log_hp1, 
                         y = median_roc_auc, 
                         group = hp3, 
                         color = hp3)) +
          geom_line() +
          facet_grid(resample ~ hp2) +
          scale_color_discrete(name = "mtry") +
          labs(title = plot_title, x = "log10 learning rate", y = "ROC AUC")
      
       print(plot_i)
       
        plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and NO RESAMPLING")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "none")) |> 
        ggplot(mapping = aes(x = log_hp1, 
                         y = median_roc_auc, 
                         group = hp3, 
                         color = hp3)) +
          geom_line() +
          facet_grid(resample ~ hp2) +
          scale_color_discrete(name = "mtry") +
          labs(title = plot_title, x = "log10 learning rate", y = "ROC AUC")
      
       print(plot_i)
      
      # hp2 x axis
      
      plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and DOWNSAMPLE")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)), # prob have to change this to factor
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "down")) |> 
        ggplot(mapping = aes(x = hp2, 
                         y = median_roc_auc, 
                         group = log_hp1, 
                         color = log_hp1)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "log10 learning rate") +
          labs(title = plot_title, x = "tree depth", y = "ROC AUC")
      
       print(plot_i)
       
      plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and UPSAMPLE")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "up")) |> 
        ggplot(mapping = aes(x = hp2, 
                         y = median_roc_auc, 
                         group = log_hp1, 
                         color = log_hp1)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "log10 learning rate") +
          labs(title = plot_title, x = "tree depth", y = "ROC AUC")
      
       print(plot_i)
       
        plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and NO RESAMPLING")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "none")) |> 
        ggplot(mapping = aes(x = hp2, 
                         y = median_roc_auc, 
                         group = log_hp1, 
                         color = log_hp1)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "log10 learning rate") +
          labs(title = plot_title, x = "tree depth", y = "ROC AUC")
      
       print(plot_i)
       
       # hp3 x axis
    plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and DOWNSAMPLE")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)), # prob have to change this to factor
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "down")) |> 
        ggplot(mapping = aes(x = hp3, 
                         y = median_roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ log_hp1) +
          scale_color_discrete(name = "tree depth") +
          labs(title = plot_title, x = "mtry", y = "ROC AUC")
      
       print(plot_i)
       
      plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and UPSAMPLE")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "up")) |> 
        ggplot(mapping = aes(x = hp3, 
                         y = median_roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ log_hp1) +
          scale_color_discrete(name = "tree depth") +
          labs(title = plot_title, x = "mtry", y = "ROC AUC")
      
       print(plot_i)
       
        plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and NO RESAMPLING")
      plot_i <- results_i |>
        mutate(log_hp1 = factor(log10(hp1)),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "none")) |> 
        ggplot(mapping = aes(x = hp3, 
                         y = median_roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ log_hp1) +
          scale_color_discrete(name = "tree depth") +
          labs(title = plot_title, x = "mtry", y = "ROC AUC")
      
       print(plot_i)
    }  
    
    # rda
    if (k == "rda") {
      
      plot_title <- str_c("Plotting RDA hyperparameters for ", i, " feature set")
      
      plot_i <- results_i |>
        mutate(hp2 = factor(hp2, ordered = TRUE),
               resample = case_when(resample == "none" ~ "none_none",
                                    TRUE ~ resample)) |> 
        separate(resample, c("resample", "under_ratio"), "_") |> 
        mutate(under_ratio = factor(under_ratio)) |> 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(under_ratio ~ resample) +
          scale_color_discrete(name = "frac identity matrix") +
          labs(title = plot_title, x = "frac common cov", y = "ROC AUC")
      
       print(plot_i)
    }  
    
     # nnet
    if (k == "nnet") {
      
      plot_title <- str_c("Plotting nnet hyperparameters for ", i, " feature set and DOWNSAMPLE by HIDDEN UNITS")
      plot_i <- results_i |>
        mutate(hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "down")) |> 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "penalty") +
          labs(title = plot_title, x = "epochs", y = "ROC AUC")
      
       print(plot_i)
       
      plot_title <- str_c("Plotting nnet hyperparameters for ", i, " feature set and UPSAMPLE by HIDDEN UNITS")
      plot_i <- results_i |>
        mutate(hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "down")) |> 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "penalty") +
          labs(title = plot_title, x = "epochs", y = "ROC AUC")
      
       print(plot_i)
  
    
    plot_title <- str_c("Plotting nnet hyperparameters for ", i, " feature set and NO RESAMPLING by HIDDEN UNITS")
      plot_i <- results_i |>
        mutate(hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) |> 
        filter(str_detect(resample, "none")) |> 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "penalty") +
          labs(title = plot_title, x = "epochs", y = "ROC AUC")
      
       print(plot_i)
    } 
  }
}
```
