---
title: "EDA on LIWC Features"
author: "Coco Yu"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---
## Code Status

## Notes

## Setup

Chunk Defaults
```{r}
knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')
```

Conflicts
```{r}
options(conflicts.policy = "depends.ok")
```

```{r}
library(tidyverse)
library(kableExtra, exclude = "group_rows")
library(lubridate)
library(janitor, include.only = c("tabyl", "clean_names"))
library(here, include.only = c("here"))

theme_set(theme_classic())
```

Source Functions
```{r}
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
```

Absolute Paths 
```{r}
path_messages <- format_path(str_c("studydata/risk/data_processed/messages"))
```

Read in Data
```{r}
liwc_ind <- read_csv(
  here(path_messages, "liwc", "liwc_features_long.csv"),
  col_types = cols()
)

liwc_cat <- read_csv(
  here(path_messages, "liwc", "liwc_features_cat.csv"),
  col_types = cols()
)
```

## EDA on liwc features generated on individual messages

sanity check: number of participants
```{r}
liwc_ind |> 
  pull(subid) |> 
  unique() |> 
  length()
```

sanity check: # of labels
```{r}
(nrow(liwc_ind))
```

sanity check: # of labels per participant (same with raw_message file)
```{r}
liwc_ind |> 
  group_by(subid) |> 
  summarize(n = n()) |> 
  summarize(
    mean = mean(n),
    sd = sd(n),
    median = median(n),
    min = min(n),
    max = max(n)
  )
```

sanity check: # of features for each prediction window
```{r}
liwc_ind |> 
  select(ends_with("3day")) |> 
  ncol()

liwc_ind |> 
  select(ends_with("1week")) |> 
  ncol()

liwc_ind |> 
  select(ends_with("3day") | ends_with("1week")) |> 
  ncol()
```

### sanity check: missingness

> **By lapse labels**

> checked the percentage of labels that lack liwc features match missingness on raw messages for 3day, and 1week prediction windows

> checked that the percentage of labels that lack liwc features in 1week prediction window and combined feature sets match (both 7.83%)

```{r}
liwc_ind |> 
  select(ends_with("3day")) |> 
  rowwise() |> 
  mutate(na = if_else(all(is.na(across(everything()))), 1, 0)) |> 
  ungroup() |>
  pull(na) |> 
  mean()

liwc_ind |> 
  select(ends_with("1week")) |> 
  rowwise() |> 
  mutate(na = if_else(all(is.na(across(everything()))), 1, 0)) |> 
  ungroup() |>
  pull(na) |> 
  mean()

liwc_ind |> 
  select(ends_with("3day") | ends_with("1week")) |>
  rowwise() |> 
  mutate(na = if_else(all(is.na(across(everything()))), 1, 0)) |> 
  ungroup() |>
  pull(na) |> 
  mean()
```

> **By subject**

> checked the percentage of missingness per participant match between raw messages and liwc features for 3day and 1week prediction windows separately

> checked that the missingness per participant in 1week prediction window and combined feature sets match

```{r}
liwc_ind |> 
  select(ends_with("3day"), subid, day_start) |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = if_else(all(is.na(across(everything()))), 1, 0)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  summarize(
    mean_mean_missing = mean(mean_missing),
    sd_mean_missing = sd(mean_missing),
    median_mean_missing = median(mean_missing),
    min_mean_missing = min(mean_missing),
    max_mean_missing = max(mean_missing)
  ) |> 
  print_kbl()

liwc_ind |> 
  select(ends_with("1week"), subid, day_start) |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = if_else(all(is.na(across(everything()))), 1, 0)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  summarize(
    mean_mean_missing = mean(mean_missing),
    sd_mean_missing = sd(mean_missing),
    median_mean_missing = median(mean_missing),
    min_mean_missing = min(mean_missing),
    max_mean_missing = max(mean_missing)
  ) |> 
  print_kbl()

liwc_ind |> 
  select(-label_num, -lapse) |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = if_else(all(is.na(across(everything()))), 1, 0)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  summarize(
    mean_mean_missing = mean(mean_missing),
    sd_mean_missing = sd(mean_missing),
    median_mean_missing = median(mean_missing),
    min_mean_missing = min(mean_missing),
    max_mean_missing = max(mean_missing)
  ) |> 
  print_kbl()
```

### EDA on features

```{r}
skim_minimal <- skimr::skim_with(numeric = skimr::sfl(mean = NULL, p25 = NULL, 
                                                      p75 = NULL, 
                                                      hist = NULL))

stats_ind <- liwc_ind |> 
  select(ends_with("3day"), ends_with("1week")) |> 
  skim_minimal() |> 
  select(-n_missing)

stats_ind |> 
  arrange(numeric.sd, numeric.p50) |> 
  print_kbl()

stats_ind |> 
  write_csv(here(path_messages, "eda", "eda_liwc_ind.csv"))
```

## EDA on concatenated messages

sanity check: number of participants
```{r}
liwc_cat |> 
  pull(subid) |> 
  unique() |> 
  length()
```

sanity check: # of labels
```{r}
(nrow(liwc_cat))
```

sanity check: # of labels per participant (same with raw_message file)
```{r}
liwc_cat |> 
  group_by(subid) |> 
  summarize(n = n()) |> 
  summarize(
    mean = mean(n),
    sd = sd(n),
    median = median(n),
    min = min(n),
    max = max(n)
  )
```

sanity check: # of features for each prediction window
```{r}
liwc_cat |> 
  select(ends_with("3day")) |> 
  ncol()

liwc_cat |> 
  select(ends_with("1week")) |> 
  ncol()

liwc_cat |> 
  select(ends_with("3day") | ends_with("1week")) |> 
  ncol()
```

### sanity check: missingness

> **By lapse labels**

> checked the percentage of labels that lack liwc features match missingness on raw messages for 3day, and 1week prediction windows

> checked that the percentage of labels that lack liwc features in 1week prediction window and combined feature sets match (both 7.83%)

```{r}
liwc_cat |> 
  select(ends_with("3day")) |> 
  rowwise() |> 
  mutate(na = if_else(all(is.na(across(everything()))), 1, 0)) |> 
  ungroup() |>
  pull(na) |> 
  mean()

liwc_cat |> 
  select(ends_with("1week")) |> 
  rowwise() |> 
  mutate(na = if_else(all(is.na(across(everything()))), 1, 0)) |> 
  ungroup() |>
  pull(na) |> 
  mean()

liwc_cat |> 
  select(ends_with("3day") | ends_with("1week")) |>
  rowwise() |> 
  mutate(na = if_else(all(is.na(across(everything()))), 1, 0)) |> 
  ungroup() |>
  pull(na) |> 
  mean()
```

> **By subject**

> checked the percentage of missingness per participant match between raw messages and liwc features for 3day and 1week prediction windows separately

> checked that the missingness per participant in 1week prediction window and combined feature sets match

```{r}
liwc_cat |> 
  select(ends_with("3day"), subid, day_start) |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = if_else(all(is.na(across(everything()))), 1, 0)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  summarize(
    mean_mean_missing = mean(mean_missing),
    sd_mean_missing = sd(mean_missing),
    median_mean_missing = median(mean_missing),
    min_mean_missing = min(mean_missing),
    max_mean_missing = max(mean_missing)
  ) |> 
  print_kbl()

liwc_cat |> 
  select(ends_with("1week"), subid, day_start) |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = if_else(all(is.na(across(everything()))), 1, 0)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  summarize(
    mean_mean_missing = mean(mean_missing),
    sd_mean_missing = sd(mean_missing),
    median_mean_missing = median(mean_missing),
    min_mean_missing = min(mean_missing),
    max_mean_missing = max(mean_missing)
  ) |> 
  print_kbl()

liwc_cat |> 
  select(-label_num, -lapse) |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = if_else(all(is.na(across(everything()))), 1, 0)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  summarize(
    mean_mean_missing = mean(mean_missing),
    sd_mean_missing = sd(mean_missing),
    median_mean_missing = median(mean_missing),
    min_mean_missing = min(mean_missing),
    max_mean_missing = max(mean_missing)
  ) |> 
  print_kbl()
```

### EDA on features

```{r}
stats_cat <- liwc_cat |> 
  select(ends_with("3day"), ends_with("1week")) |> 
  skim_minimal() |> 
  select(-n_missing)

stats_cat |> 
  arrange(numeric.sd, numeric.p50) |> 
  print_kbl()

stats_cat |> 
  write_csv(here(path_messages, "eda", "eda_liwc_cat.csv"))
```
