---
title: "Processes `r params$model` models of training jobs from CHTC for `r params$study` study with version `r params$version` using `r params$cv` CV"
author: "John Curtin & Kendra Wyant"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  study: "messages"
  version: "v17"
  cv: "kfold"
  algorithms: "random_forest"   # "all" or name of specific algorithm
  model: "meta" # "main" or "baseline"
editor_options: 
  chunk_output_type: console
---

### Code Status

In use for both kfold and nested cv, including use of batches

This script aggregates all results/metrics for a batch or batches of jobs that train all model configurations for a specific outcome/label window.

### Set Up Environment

```{r}
study <- params$study
version <- params$version
cv <- params$cv
algorithms <- params$algorithms
model <- params$model
```

Handle conflicts

```{r}
#| message: false
#| warning: false

options(conflicts.policy = "depends.ok")
```

Packages for script

```{r}
#| message: false
#| warning: false

library(tidyverse)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
theme_set(theme_classic()) 
```

Absolute paths

```{r}
#| message: false

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_input <- format_path(str_c("risk/chtc/", study))
path_models <- format_path(str_c("risk/models/", study))
```

Chunk Defaults

```{r}
#| include: false

knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')

options(tibble.width = Inf)
options(tibble.print_max = Inf)
```

### Read results.csv files

Set up object for results

```{r}
results_all <- NULL
```

Get batch_names

```{r}
batch_names <- list.dirs(path_input, full.names = FALSE, recursive = FALSE)
  
if (algorithms == "all") {
  batch_names <- batch_names[str_detect(batch_names, "train") & 
                               str_detect(batch_names, cv) &
                               str_detect(batch_names, version) &
                               str_detect(batch_names)]
  if (model != "main") {
    batch_names <- batch_names[str_detect(batch_names, model)]
  }
} else {
    batch_names <- batch_names[str_detect(batch_names, "train") & 
                               str_detect(batch_names, cv) &
                               str_detect(batch_names, version) &
                               str_detect(batch_names, algorithms)]
  if (model != "main") {
    batch_names <- batch_names[str_detect(batch_names, model)]
  }
}  
 
batch_names
```

Loop over batch_names to read in files and perform checks

```{r}

for (batch_name in batch_names) {
  message("Processing Batch: ", batch_name)
  
  # read in configs
  configs <- read_csv(here::here(path_input, batch_name, "input", "configs.csv"), 
                      show_col_types = FALSE)
  (n_configs <- nrow(configs))
  
  # read in results
  results_batch <- read_csv(here::here(path_input, batch_name, "output", 
                                      "batch_results.csv"), 
                            show_col_types = FALSE)
  (n_results_batch <- nrow(results_batch))
  
  # Check counts of results files
  if (str_detect(batch_name, "glmnet")) {
    source(here::here(path_input, batch_name, "input/training_controls.R"))
    check <- (n_configs) == (n_results_batch)/hp2_glmnet_out
  } else {
    check <- n_configs == n_results_batch
  }  
  if (!check) {
    stop(n_configs, " configs != ", n_results_batch, " results files!")
  } else {
    message(n_results_batch, " results files detected.  Correct!")
  }
  
  # Check col count
  if (!(ncol(results_batch) == 17)) {
    stop(ncol(results_batch), " columns != 17")
  } else {
    message(ncol(results_batch), " columns detected.  Correct!\n")
  }


  # Add batch to all metrics
  results_all <- results_all %>% 
    bind_rows(results_batch)
}
```

### Wrap up processing of raw metrics

Remove duplicate rows (e.g., same hyper-parameters across multiple batches)

```{r}
nrow(results_all)

results_all <- results_all |> 
  distinct(split_num, outer_split_num, inner_split_num, algorithm, feature_set,
           hp1, hp2, hp3, resample, .keep_all = TRUE)

nrow(results_all)
```

Final checks across all batches

```{r}
  results_all %>% tab(split_num) %>% print()
  results_all %>% tab(outer_split_num) %>% print()
  results_all %>% tab(inner_split_num) %>% print()
  results_all %>% tab(algorithm) %>% print()
  results_all %>% tab(feature_set) %>% print()
  results_all %>% tab(hp1) %>% print()
  results_all %>% tab(hp2) %>% print()
  results_all %>% tab(hp3) %>% print()
  results_all %>% tab(resample) %>% print()
```

Save raw metrics file

```{r}
results_all %>% 
  # arrange(split_num, outer_split_num, inner_split_num, algorithm, resample
  write_csv(here::here(path_models, str_c("inner_metrics_", 
                                            window, "_", lead, "_", version, "_", 
                                            cv, "_", model, ".csv")))
```

### Median metrics across inner folds for model configurations

Inner loop performance of best config.
This median performance for each configuration over inner x outer folds (e.g., 300 folds for 1x10 inner and 3x10 outer). It is what we would get (essentially) if we just did simple k-fold but with LOTs of folds

```{r}
metrics_avg <- results_all %>% 
  group_by(algorithm, feature_set, hp1, hp2, hp3, resample) %>% 
   summarize(roc_auc = median(roc_auc),
             n_jobs = n(), .groups = "drop") %>% 
  relocate(n_jobs) %>% 
  arrange(desc(roc_auc)) |> 
  ungroup()

metrics_avg |> 
  group_by(algorithm) |> 
  slice(1) |> 
  arrange(desc(roc_auc)) |> 
  print_kbl()

metrics_avg |> 
  slice(1:50) |> 
  print_kbl()

best_config <- metrics_avg |> 
  slice(1) |> 
  print()
```

Performance metric plot across all inner folds

```{r}
results_all |> 
  filter(algorithm == best_config$algorithm,
         feature_set == best_config$feature_set,
         hp1 == best_config$hp1,
         hp2 == best_config$hp2,
         hp3 == best_config$hp3,
         resample == best_config$resample) |> 
  ggplot(aes(x = roc_auc)) +
  geom_histogram(bins = 10)
```


### feature sets for good vs poor splits
```{r}
feats_folds <- results_all |> 
  filter(algorithm == best_config$algorithm,
         feature_set == best_config$feature_set,
         hp1 == best_config$hp1,
         hp2 == best_config$hp2,
         hp3 == best_config$hp3,
         resample == best_config$resample) |> 
  select(split_num, meta, baseline, roc_auc)
```

22 folds over .7 auroc

- most had the minimum meta features allowed (40)
```{r}
feats_folds |> 
  filter(roc_auc > .7) |> 
  mutate(n_meta = lengths(str_split(meta, ","))) |> 
  pull(n_meta)

feats_folds |> 
  filter(roc_auc > .7) |> 
  mutate(n_meta = lengths(str_split(meta, ","))) |> 
  pull(n_meta) |> 
  mean()
```

```{r}
feats_folds |> 
  filter(roc_auc < .6) |> 
  mutate(n_meta = lengths(str_split(meta, ","))) |> 
  pull(n_meta)

feats_folds |> 
  filter(roc_auc < .6) |> 
  mutate(n_meta = lengths(str_split(meta, ","))) |> 
  pull(n_meta) |> 
  mean()
```

```{r}
feats_folds |> 
  filter(roc_auc >= .6 & roc_auc <= .7) |> 
  mutate(n_meta = lengths(str_split(meta, ","))) |> 
  pull(n_meta) 

feats_folds |> 
  filter(roc_auc >= .6 & roc_auc <= .7) |> 
  mutate(n_meta = lengths(str_split(meta, ","))) |> 
  pull(n_meta) |> 
  mean()
```

Baseline feature range

```{r}
feats_folds |> 
  filter(roc_auc > .7) |> 
  mutate(n_base = lengths(str_split(baseline, ","))) |> 
  pull(n_base)

feats_folds |> 
  filter(roc_auc > .7) |> 
  mutate(n_base = lengths(str_split(baseline, ","))) |> 
  pull(n_base) |> 
  mean()
```

```{r}
feats_folds |> 
  filter(roc_auc < .6) |> 
  mutate(n_base = lengths(str_split(baseline, ","))) |> 
  pull(n_base)

feats_folds |> 
  filter(roc_auc < .6) |> 
  mutate(n_base = lengths(str_split(baseline, ","))) |> 
  pull(n_base) |> 
  mean()
```


```{r}
feats_folds |> 
  filter(roc_auc >= .6 & roc_auc <= .7) |> 
  mutate(n_base = lengths(str_split(baseline, ","))) |> 
  pull(n_base) 

feats_folds |> 
  filter(roc_auc >= .6 & roc_auc <= .7) |> 
  mutate(n_base = lengths(str_split(baseline, ","))) |> 
  pull(n_base) |> 
  mean()
```

Look at unique baseline features in good folds (auroc > .7)
```{r}
feats_folds |> 
  filter(roc_auc > .7) |> 
  select(-meta) |> 
  separate_rows(baseline, sep = ",") 
```

and bad folds
```{r}
feats_folds |> 
  filter(roc_auc < .6) |> 
  select(-meta) |> 
  separate_rows(baseline, sep = ",") 
```

Number of unique baseline items across all splits is 103
```{r}
feats_folds |> 
  select(-meta) |> 
  separate_rows(baseline, sep = ",") |> 
  mutate(baseline = str_remove(baseline, " ")) |> 
  count(baseline) |> 
  arrange(desc(n))
```


Number of unique meta items across all splits 
```{r}
feats_folds |> 
  select(-baseline) |> 
  separate_rows(meta, sep = ",") |> 
  mutate(meta = str_remove(meta, "^ "),
         meta = str_remove(meta, "`")) |> 
  count(meta) |> 
  arrange(desc(n))
```


### Plot hyperparameters

```{r}
algorithms <- unique(metrics_avg$algorithm) 
feature_sets <- unique(metrics_avg$feature_set) 

for (k in algorithms) {
  
  results_k <- metrics_avg %>% 
      filter(algorithm == k)
  
  for (i in feature_sets) {
  
    results_i <- results_k %>% 
      filter(feature_set == i)
    
    
    # glmnet
    if (k == "glmnet") {
  
      plot_title <- str_c("Plotting glmnet hyperparameters for ", i, " feature set")
  
  
      plot_i <- results_i %>%
        mutate(hp1 = factor(hp1, ordered = TRUE),
               resample = case_when(resample == "none" ~ "none_none",
                                    TRUE ~ resample)) %>% 
        separate(resample, c("resample", "under_ratio"), "_") %>% 
        mutate(under_ratio = factor(under_ratio)) %>% 
        ggplot(mapping = aes(x = log(hp2), 
                         y = roc_auc, 
                         group = hp1, 
                         color = hp1)) +
          geom_line() +
          facet_grid(under_ratio ~ resample) +
          scale_color_discrete(name = "mixture (alpha)") +
          labs(title = plot_title, x = "penalty (lambda)", y = "ROC AUC")
  
      print(plot_i)
    }

    
    # XGBoost
    if (k == "xgboost") {
      
      plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and DOWNSAMPLE")
      plot_i <- results_i %>%
        mutate(log_hp1 = log10(hp1),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) %>% 
        filter(str_detect(resample, "down")) %>% 
        ggplot(mapping = aes(x = log_hp1, 
                         y = roc_auc, 
                         group = hp3, 
                         color = hp3)) +
          geom_line() +
          facet_grid(resample ~ hp2) +
          scale_color_discrete(name = "mtry") +
          labs(title = plot_title, x = "log10 learning rate", y = "ROC AUC")
      
       print(plot_i)
       
      plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and UPSAMPLE")
      plot_i <- results_i %>%
        mutate(log_hp1 = log10(hp1),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) %>% 
        filter(str_detect(resample, "up")) %>% 
        ggplot(mapping = aes(x = log_hp1, 
                         y = roc_auc, 
                         group = hp3, 
                         color = hp3)) +
          geom_line() +
          facet_grid(resample ~ hp2) +
          scale_color_discrete(name = "mtry") +
          labs(title = plot_title, x = "log10 learning rate", y = "ROC AUC")
      
       print(plot_i)
       
        plot_title <- str_c("Plotting XGBoost hyperparameters for ", i, " feature set and NO RESAMPLING")
      plot_i <- results_i %>%
        mutate(log_hp1 = log10(hp1),
               hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) %>% 
        filter(str_detect(resample, "none")) %>% 
        ggplot(mapping = aes(x = log_hp1, 
                         y = roc_auc, 
                         group = hp3, 
                         color = hp3)) +
          geom_line() +
          facet_grid(resample ~ hp2) +
          scale_color_discrete(name = "mtry") +
          labs(title = plot_title, x = "log10 learning rate", y = "ROC AUC")
      
       print(plot_i)
    }  
    
    # rda
    if (k == "rda") {
      
      plot_title <- str_c("Plotting RDA hyperparameters for ", i, " feature set")
      
      plot_i <- results_i %>%
        mutate(hp2 = factor(hp2, ordered = TRUE),
               resample = case_when(resample == "none" ~ "none_none",
                                    TRUE ~ resample)) %>% 
        separate(resample, c("resample", "under_ratio"), "_") %>% 
        mutate(under_ratio = factor(under_ratio)) %>% 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(under_ratio ~ resample) +
          scale_color_discrete(name = "frac identity matrix") +
          labs(title = plot_title, x = "frac common cov", y = "ROC AUC")
      
       print(plot_i)
    }  
    
     # nnet
    if (k == "nnet") {
      
      plot_title <- str_c("Plotting nnet hyperparameters for ", i, " feature set and DOWNSAMPLE by NNET")
      plot_i <- results_i %>%
        mutate(hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) %>% 
        filter(str_detect(resample, "down")) %>% 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "penalty") +
          labs(title = plot_title, x = "epochs", y = "ROC AUC")
      
       print(plot_i)
       
      plot_title <- str_c("Plotting nnet hyperparameters for ", i, " feature set and UPSAMPLE by NNET")
      plot_i <- results_i %>%
        mutate(hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) %>% 
        filter(str_detect(resample, "down")) %>% 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "penalty") +
          labs(title = plot_title, x = "epochs", y = "ROC AUC")
      
       print(plot_i)
  
    
    plot_title <- str_c("Plotting nnet hyperparameters for ", i, " feature set and NO RESAMPLING by NNET")
      plot_i <- results_i %>%
        mutate(hp2 = factor(hp2, ordered = TRUE),
               hp3 = factor(hp3, ordered = TRUE),
               resample = factor(resample)) %>% 
        filter(str_detect(resample, "none")) %>% 
        ggplot(mapping = aes(x = hp1, 
                         y = roc_auc, 
                         group = hp2, 
                         color = hp2)) +
          geom_line() +
          facet_grid(resample ~ hp3) +
          scale_color_discrete(name = "penalty") +
          labs(title = plot_title, x = "epochs", y = "ROC AUC")
      
       print(plot_i)
    } 
  }
}
```
