---
title: "EDA on Raw Messages"
author: "Coco Yu"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Code Status

## Notes

## Setup

Chunk Defaults
```{r}
knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')
theme_set(theme_classic())
```

Conflicts
```{r}
options(conflicts.policy = "depends.ok")
```

```{r}
library(tidyverse)
library(kableExtra, exclude = "group_rows")
library(lubridate)
library(janitor, include.only = c("tabyl", "clean_names"))
library(here, include.only = c("here"))
```

Source Functions
```{r}
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
```

Absolute Paths 
```{r}
path_shared <- format_path(str_c("studydata/risk/data_processed/shared"))
path_messages <- format_path(str_c("studydata/risk/data_processed/messages"))
```

Read in Data 
```{r}
# raw messages that have been cleaned
raw_data <- read_csv(
    here(path_messages, "cleaned_messages.csv"),
    col_types = cols(
        subid = "d", address = "c", contact_type = "f", date = "T", type = "f", 
        body = "c", msg_type = "f", phone_type = "f"
    )
)

# individual messages mapped on to prediction windows
pred_3day <- read_csv(
    here(path_messages, "predictors/pred_3day.csv"),
    col_types = cols(
        subid = "d", day_start = "T", day_end = "T", lapse = "f", pred_onset = "T",
        address = "c", contact_type = "f", date = "T", type = "f", body = "c",
        msg_type = "f", phone_type = "f"
    )
)
pred_1week <- read_csv(
    here(path_messages, "predictors/pred_1week.csv"),
    col_types = cols(
        subid = "d", day_start = "T", day_end = "T", lapse = "f", pred_onset = "T",
        address = "c", contact_type = "f", date = "T", type = "f", body = "c",
        msg_type = "f", phone_type = "f"
    )
)

# concatenated text messages mapped on to prediction windows
cat_3day <- read_csv(
    here(path_messages, "predictors/doc_3day.csv"),
    col_types = cols(
        subid = "d", pred_onset = "T", body = "c", lapse = "f"
    )
)

cat_1week <- read_csv(
    here(path_messages, "predictors/doc_1week.csv"),
    col_types = cols(
        subid = "d", pred_onset = "T", body = "c", lapse = "f"
    )
)

# labels
labels <- read_csv(here(path_messages, "lapses.csv"), col_types = cols()) |> 
  mutate(day_start = as_datetime(day_start, tz = "America/Chicago"),
         day_end = as_datetime(day_end, tz = "America/Chicago"))
```

## EDA on raw messages

### Message counts per participant

**Summary Table**

> Conclusion 1: The total number of messages in the dataset is 313549.

> Conclusion 2: 

```{r}
(nrow(raw_data))

raw_data |> 
  group_by(subid) |> 
  summarise(n = n()) |>
  summarise(
    mean = mean(n),
    sd = sd(n),
    median = median(n),
    min = min(n),
    max = max(n)
  )
```

**Histogram on # of messages per participant**
```{r}
raw_data |> 
    group_by(subid) |> 
    summarise(n = n()) |> 
    ggplot(aes(x = n)) + 
    geom_histogram(bins = 50, alpha = .6) +
    labs(x = "# of messages per participant")
```

### Message Length per participant

**Summary Table**

> Conclusion 1: The average message length is 10.4 (range = 1 - 1266)

> Conclusion 2: On average, each participant has a mean message length of 10.9 (range = 5.25 - 38.91) and a median message length of 7.14 (4 - 19).

```{r}
# mean/median message length across all messages
raw_data |> 
  mutate(text_length = lengths(strsplit(raw_data$body, "\\s+")))  |> 
  summarize(
    mean_length = mean(text_length),
    median_length = median(text_length),
    min_length = min(text_length),
    max_length = max(text_length)
  )

# summary statistics for mean/median message length per participant
raw_data |> 
  mutate(text_length = lengths(strsplit(raw_data$body, "\\s+")))  |> 
  group_by(subid) |> 
  summarize(
    mean_length = mean(text_length),
    median_length = median(text_length),
    min_length = min(text_length),
    max_length = max(text_length)
  ) |> 
  summarize(
    mean_mean_length = mean(mean_length),
    median_mean_length = median(mean_length),
    min_mean_length = min(mean_length),
    max_mean_length = max(mean_length),
    mean_median_length = mean(median_length),
    median_median_length = median(median_length),
    min_median_length = min(median_length),
    max_median_length = max(median_length)
  ) |> 
  print_kbl()
```

**Histogram on mean/median message length per participant**
```{r}
raw_data |> 
  mutate(text_length = lengths(strsplit(raw_data$body, "\\s+")))  |> 
  group_by(subid) |> 
  summarize(mean_length = mean(text_length)) |> 
  ggplot(aes(x = mean_length)) + 
  geom_histogram(binwidth = 1, alpha = .6) +
  labs(x = "mean message length per participant")

raw_data |> 
  mutate(text_length = lengths(strsplit(raw_data$body, "\\s+")))  |> 
  group_by(subid) |> 
  summarize(median_length = median(text_length)) |> 
  ggplot(aes(x = median_length)) + 
  geom_histogram(binwidth = 1, alpha = .6) +
  labs(x = "median message length per participant")
```

## EDA on prediction windows

prep: add in lapses that don't have any predictors
```{r}
pred_3day <- pred_3day |> 
    full_join(
        labels,
        by = c("subid", "day_start", "day_end", "lapse")
    )

pred_1week <- pred_1week |> 
    full_join(
        labels,
        by = c("subid", "day_start", "day_end", "lapse")
    )

cat_3day <- cat_3day |> 
    full_join(
        labels |> 
            mutate(pred_onset = day_start - days(3)),
        by = c("subid", "pred_onset", "lapse")
    )

cat_1week <- cat_1week |> 
    full_join(
        labels |> 
            mutate(pred_onset = day_start - days(3)),
        by = c("subid", "pred_onset", "lapse")
    )
```

sanity check
```{r}
nrow(
    pred_3day |> 
        group_by(subid, day_start) |> 
        summarize(na = any(is.na(body))) |> nrow()
        pull(na) |> 
        sum()
) == 
nrow(
    cat_3day |> 
        filter(is.na(body))
)
```