---
title: "EDA on Raw Messages"
author: "Coco Yu"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Code Status

## Notes

## Setup

Chunk Defaults
```{r}
knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')
```

Conflicts
```{r}
options(conflicts.policy = "depends.ok")
```

```{r}
library(tidyverse)
library(kableExtra, exclude = "group_rows")
library(lubridate)
library(janitor, include.only = c("tabyl", "clean_names"))
library(here, include.only = c("here"))

theme_set(theme_classic())
```

Source Functions
```{r}
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
```

Absolute Paths 
```{r}
# path_shared <- format_path(str_c("studydata/risk/data_processed/shared"))
path_messages <- format_path(str_c("studydata/risk/data_processed/messages"))
```

Read in Data 
```{r}

raw_data <- read_csv(here(path_messages, "eda", "eda_raw.csv")) |> glimpse()
pred_3day <- read_csv(here(path_messages, "eda", "eda_3day.csv")) |> glimpse()
pred_1week <- read_csv(here(path_messages, "eda", "eda_1week.csv")) |> glimpse()

labels <- read_csv(here(path_messages, "lapses.csv"), col_types = cols()) |> 
  mutate(day_start = as_datetime(day_start, tz = "America/Chicago"),
         day_end = as_datetime(day_end, tz = "America/Chicago"))
```

## EDA on lapse labels

```{r}
#| label: fig-lapse_count
#| fig-cap: "Distribution of Lapse Label Counts"

labels |> 
  group_by(subid) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = n)) + 
  geom_histogram(fill = "#97C4C5", bins = 70) +
  labs(y = "subject counts", x = "# of lapses")
```

## EDA on raw messages

### Message counts per participant

**Histogram on # of messages per participant**
```{r}
#| label: fig-message_count
#| fig-cap: "Distribution of Message Counts"
#| fig-width: 3.5
#| fig-height: 2.5
plt1_raw <- raw_data |> 
  group_by(subid) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = n)) + 
  geom_histogram(bins = 50, fill = "#97C4C5") +
  labs(x = "# of messages", y = "subject counts",
       caption = str_wrap("(a) Distribution of Message Counts by Subject", 
                          width = 70)) +
  theme(plot.caption = element_text(hjust = 0.5))
```

### Message Length per participant

**Histogram on mean/median message length per participant**
```{r}
#| label: fig-message_length
#| fig-cap: "Distribution of Mean Message Length"
#| fig-width: 3.5
#| fig-height: 2.5
plt2_raw <- raw_data |> 
  group_by(subid) |> 
  summarize(mean_length = mean(text_length)) |> 
  ggplot(aes(x = mean_length)) + 
  geom_histogram(binwidth = 1, fill = "#97C4C5") +
  labs(x = "mean message length", y = "subject counts", 
       caption = str_wrap("(b) Distribution of Mean Message Length by Subject"),
       width = 70) +
  theme(plot.caption = element_text(hjust = 0.5))
```

```{r}
raw_data |> 
  group_by(subid) |> 
  summarize(median_length = median(text_length)) |> 
  ggplot(aes(x = median_length)) + 
  geom_histogram(binwidth = 1, fill = "#97C4C5") +
  labs(x = "median message length", y = "subject counts")
```

```{r}
#| label: fig-raw
#| fig-cap: "Sample Characteristics of Raw Messages"

ggpubr::ggarrange(plt1_raw, plt2_raw)
```

## EDA on prediction windows

### prediction window: 3-day

**Distribution by label**
```{r}
plt1_3day <- pred_3day |> 
  group_by(subid, day_start) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = n)) + 
  geom_histogram(fill = "#97C4C5") +
  labs(x = "# of messages", y = "lapse counts", 
       caption = str_wrap("(a) Distribution of Message Counts"), width = 70) +
  theme(plot.caption = element_text(hjust = 0.5)) 
```

**Distribution by participant**
```{r}
plt2_3day <- pred_3day |> 
  group_by(subid, day_start) |> 
  summarize(
    n_messages = sum(!na)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_messages = mean(n_messages)
  ) |>
  ggplot(aes(x = mean_messages)) +
  geom_histogram(fill = "#97C4C5") +
  labs(x = "average # of messages per label", y = "subject counts",
       caption = str_wrap("(b) Distribution of Message Counts by Subject",
                          width = 70)) +
  theme(plot.caption = element_text(hjust = 0.5)) 
```

```{r}
plt3_3day <- pred_3day |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = sum(na) / n()
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  ggplot(aes(x = mean_missing)) +
  geom_histogram(fill = "#97C4C5") +
  labs(x = "percentage of missingness", y = "subject counts",
       caption = str_wrap("(c) Distribution of Missingness by Subject", 
                          width = 70)) +
  theme(plot.caption = element_text(hjust = 0.5)) 
```

```{r}
#| label: fig-3day
#| fig-cap: "Sample Characteristics in the 3-Day Prediction Window"
ggpubr::ggarrange(
  ggpubr::ggarrange(NULL, plt1_3day, NULL, ncol = 3, widths = c(1, 1, 1)), 
  ggpubr::ggarrange(plt2_3day, plt3_3day, ncol = 2, nrow = 1),  
  nrow = 2, ncol = 1, heights = c(1, 1) 
)

```

### prediction window: 1 week

**Distribution by label**
```{r}
plt1_1week <- pred_1week |> 
  group_by(subid, day_start) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = n)) + 
  geom_histogram(fill = "#97C4C5") +
  labs(x = "# of messages", y = "lapse counts", 
       caption = str_wrap("(a) Distribution of Message Counts"), width = 70) +
  theme(plot.caption = element_text(hjust = 0.5)) 
```

```{r}
plt2_1week <- pred_1week |> 
  group_by(subid, day_start) |> 
  summarize(
    n_messages = sum(!na)
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_messages = mean(n_messages)
  ) |>
  ggplot(aes(x = mean_messages)) +
  geom_histogram(fill = "#97C4C5") +
  labs(x = "average # of messages per label", y = "subject counts",
       caption = str_wrap("(b) Distribution of Message Counts by Subject",
                          width = 70)) +
  theme(plot.caption = element_text(hjust = 0.5)) 
```

```{r}
plt3_1week <- pred_1week |> 
  group_by(subid, day_start) |> 
  summarize(
    missingness = sum(na) / n()
  ) |> 
  group_by(subid) |> 
  summarize(
    mean_missing = mean(missingness)
  ) |> 
  ggplot(aes(x = mean_missing)) +
  geom_histogram(fill = "#97C4C5") +
  labs(x = "percentage of missingness", y = "subject counts",
       caption = str_wrap("(c) Distribution of Missingness by Subject", 
                          width = 70)) +
  theme(plot.caption = element_text(hjust = 0.5)) 
```

```{r}
#| label: fig-1week
#| fig-cap: "Sample Characteristics in the 1-Week Prediction Window"
ggpubr::ggarrange(
  ggpubr::ggarrange(NULL, plt1_1week, NULL, ncol = 3, widths = c(1, 1, 1)), 
  ggpubr::ggarrange(plt2_1week, plt3_1week, ncol = 2, nrow = 1),  
  nrow = 2, ncol = 1, heights = c(1, 1) 
)

```