---
title: "Make LIWC features"
author: "Coco Yu"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Notes

- Coco removed all OBJs and emojis from the raw text

- Emoji category was removed from LIWC dictionary

- Around 82% of feature q_10 score and median score overlap

## Setup

Chunk Defaults
```{r}
knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')
```

Conflicts
```{r}
options(conflicts.policy = "depends.ok")
```

```{r}
library(tidyverse)
library(kableExtra, exclude = "group_rows")
library(lubridate)
library(janitor, include.only = c("tabyl", "clean_names"))
library(here, include.only = c("here"))
```

Source Functions
```{r}
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
```

Absolute Paths 
```{r}
path_messages <- format_path(str_c("studydata/risk/data_processed/messages"))
```


## Clean LIWC Results and Engineer Features

read in lapse labels
```{r}
labels <- read_csv(here(path_messages, "lapses.csv"), col_types = cols()) |> 
  mutate(day_start = as_datetime(day_start, tz = "America/Chicago"),
         day_end = as_datetime(day_end, tz = "America/Chicago"))
```

```{r}
clean_features <- function(window){

  # read in data
  data <- read_csv(
    here(path_messages, str_c("liwc/liwc_", window, ".csv")),
    col_types = cols()
  ) |> 
  janitor::clean_names()

  # remove unnecessary columns
  data <- data |> 
    select(
      -day_end, -pred_onset, -address, -contact_type, -date, -type, -msg_type,
      -phone_type, -column_id, -segment, -emoji
    )
  
  # normalize liwc scores
  data <- data |> 
    mutate(across(
      -c(subid, day_start, lapse, text, wc, wps, analytic, clout, authentic, tone),
      ~ (. * wc / 100) / sqrt(wc)
    ))

  # create new columns
  data <- data |> 
    group_by(subid, day_start, lapse) |> 
    summarize(across(
      -text,
      list(
        median = ~ median(., na.rm = TRUE),
        q_10 = ~ quantile(., 0.1, na.rm = TRUE),
        q_90 = ~ quantile(., 0.9, na.rm = TRUE)
      ),
      .names = str_c("{.col}_{.fn}_", window)
    ),
    .groups = "drop")
  
  # merge with lapse labels
  data <- data |> 
    full_join(
      labels,
      by = c("subid", "day_start", "lapse")
    ) |> 
    select(-day_end)
}
```

create dataframes
```{r}
liwc_3day <- clean_features("3day")
liwc_1week <- clean_features("1week")

liwc_combined <- liwc_3day |> 
  full_join(liwc_1week, by = c("subid", "day_start", "lapse"))
```

## EDA

look at means, median, sd and min/max

```{r}
missing_3day <- liwc_3day |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  pull(missing) |> 
  mean()

missing_1week <- liwc_1week |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  pull(missing) |> 
  mean()

missing_combined <- liwc_combined |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  pull(missing) |> 
  mean()

liwc_3day |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  group_by(subid) |> 
  summarise(
    miss = sum(missing) / n()
  ) |> 
  pull(miss) |> 
  mean()
```

```{r}
liwc_3day |> 
  select(
    -subid, -day_start, -lapse, 
    -contains(c("analytic", "tone", "clout", "authentic"))) |> 
  summarise(
    across(everything(), ~ mean(is.na(.x)))
  ) |> 
  as.matrix() |> 
  t() |> 
  as_tibble(rownames = "variable") |> 
  filter(round(V1) != round(missing_3day))

liwc_1week|> 
  select(
    -subid, -day_start, -lapse, 
    -contains(c("analytic", "tone", "clout", "authentic"))) |> 
  summarise(
    across(everything(), ~ mean(is.na(.x)))
  ) |> 
  as.matrix() |> 
  t() |> 
  as_tibble(rownames = "variable") |> 
  filter(round(V1) != round(missing_1week))
```

```{r}
liwc_3day |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  group_by(subid) |> 
  summarize(mean_missing = mean(missing))

liwc_1week |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  group_by(subid) |> 
  summarize(mean_missing = mean(missing))

liwc_1week |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  group_by(subid) |> 
  summarize(mean_missing = mean(missing))
```

```{r}
liwc_3day |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  group_by(subid) |> 
  summarize(n_missing = sum(missing) / n()) |> 
  pull(n_missing) |> 
  hist()

liwc_1week |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  group_by(subid) |> 
  summarize(n_missing = sum(missing) / n()) |> 
  pull(n_missing) |> 
  hist()

liwc_combined |> 
  rowwise() |> 
  mutate(
    missing = if_else(all(is.na(c_across(-c(subid, day_start, lapse)))), 1, 0)
  ) |> 
  ungroup() |> 
  group_by(subid) |> 
  summarize(n_missing = sum(missing) / n()) |> 
  pull(n_missing) |> 
  hist()
```

```{r}
liwc_3day |> 
  select(-subid, -day_start, -lapse) |> 
  skim_some()

liwc_1week |> 
  select(-subid, -day_start, -lapse) |> 
  skim_some()

liwc_combined |> 
  select(-subid, -day_start, -lapse) |> 
  skim_some()
```

check q10 lower than q90

```{r}
check <- read_csv(
    here(path_messages, str_c("liwc/liwc_", "3day", ".csv")),
    col_types = cols()
  ) |> 
  janitor::clean_names()
```

```{r}
check |> pull(adverb) |> hist()


check |> 
  mutate(adverb = adverb / 100 * wc / sqrt(wc)) |> 
  # group_by(subid, day_start) |> 
  # summarize(median = median(adverb)) |> 
  # pull(median) |> 
  pull(adverb) |> 
  hist()
```

make scatter plots to see relationship between normalized on wc, normalized on sqrt(wc), and wc

```{r}
# remove unnecessary columns
check <- check |> 
  select(
    -day_end, -pred_onset, -address, -contact_type, -date, -type, -msg_type,
    -phone_type, -column_id, -segment, -emoji
  )

check <- check |> 
  pivot_longer(
    cols = -c(subid, day_start, lapse, text, wc, wps),
    names_to = "feature",
    values_to = "norm_wc"
  ) |> 
  mutate(
    count = norm_wc / 100 * wc,
    norm_sqrt = norm_wc / 100 * wc / sqrt(wc),
    norm_log = norm_wc / 100 * wc / log(wc + 1)
  ) 

check_agg <- check |> 
  group_by(subid, day_start, feature) |> 
  summarise(
    q_1 = quantile(norm_sqrt, 0.01, na.rm = TRUE),
    q_5 = quantile(norm_sqrt, 0.05, na.rm = TRUE),
    q_10 = quantile(norm_sqrt, 0.1, na.rm = TRUE),
    q_50 = median(norm_sqrt, na.rm = TRUE),
    q_90 = quantile(norm_sqrt, 0.9, na.rm = TRUE),
    q_95 = quantile(norm_sqrt, 0.95, na.rm = TRUE)
  )

(check_agg |> filter(q_10 == q_50) |> nrow()) / nrow(check_agg)
(check_agg |> filter(q_5 == q_50) |> nrow()) / nrow(check_agg)
(check_agg |> filter(q_1 == q_50) |> nrow()) / nrow(check_agg)
(check_agg |> filter(q_90 == q_50) |> nrow()) / nrow(check_agg)
(check_agg |> filter(q_95 == q_50) |> nrow()) / nrow(check_agg)
  
check |> 
  ggplot(aes(x = norm_sqrt)) +
  geom_histogram() +
  facet_wrap(~feature, scales = "free") +
  theme_classic()

```


