---
title: "Clean up indiviudal messages"
author: "Coco Yu"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Code Status

## Setup

Chunk Defaults
```{r}
knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')
```

Conflicts
```{r}
options(conflicts.policy = "depends.ok")
```

```{r}
library(tidyverse)
library(kableExtra, exclude = "group_rows")
library(lubridate)
library(janitor, include.only = c("tabyl", "clean_names"))
library(here, include.only = c("here"))
```

Source Functions
```{r}
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
```

Absolute Paths 
```{r}
path_shared <- format_path(str_c("studydata/risk/data_processed/shared"))
path_messages <- format_path(str_c("studydata/risk/data_processed/messages"))
```

## Read in Data

```{r}
data <- read_csv(here(path_shared, "sms.csv"), 
                 col_types = cols(
                   subid = "d", start_study = "T", end_study = "T",
                   address = "c", date = "T", type = "f", body = "c",
                   msg_type = "f", phone_type = "f"
                 )) |> 
  mutate(date = as_datetime(date, tz = "America/Chicago"),
         start_study = as_datetime(start_study, tz = "America/Chicago"),
         end_study = as_datetime(end_study, tz = "America/Chicago")) |> 
  glimpse()
```

## Data Cleaning

### Add Contact Information

read in contact data
```{r}
contacts <- read_csv(here(path_shared, "contacts.csv"),
                     col_types = cols(.default = col_character(),
                                      subid = "d"))
```

check format of address in contacts.csv
```{r}
nchar(contacts$phone_number) |> 
  tabyl()

contacts |> 
  mutate(
    digits = nchar(phone_number)
  ) |> 
  group_by(digits) |> 
  slice_head(n = 20) |> 
  print_kbl()
```

make address format consistent across contacts and data
```{r}
contacts <- contacts |> 
  mutate(
    phone_number = if_else(
      nchar(phone_number) == 10,
      paste0("1", phone_number),
      phone_number
    )
  )
```

combine data and contacts
```{r}
data <- contacts |> 
  group_by(subid) |> 
  summarize(freq_contacts = list(phone_number)) |>
  right_join(data, by = "subid") |> 
  glimpse()
```

create new column freq_contacts: y/n
```{r}
data <- data |> 
  mutate(
    freq_contacts = if_else(
      address %in% unlist(freq_contacts),
      "yes", "no"
    ),
    freq_contacts = factor(freq_contacts)
  ) |> 
  glimpse()
```

### Add lapse information

read in lapse labels
```{r}
labels <- read_csv(here(path_messages, "lapses.csv"), col_types = cols()) |> 
  mutate(lapse_onset = as_datetime(lapse_onset, tz = "America/Chicago")) |> 
  glimpse()
```

create prediction window
```{r}
pred_1day <- labels |> 
  mutate(pred_1day = lapse_onset - days(1)) |> 
  inner_join(data, by = "subid") |>
  filter(date >= pred_1day & date <= lapse_onset) |>
  glimpse()

pred_3day <- labels |> 
  mutate(pred_3day = lapse_onset - days(3)) |> 
  inner_join(data, by = "subid") |>
  filter(date >= pred_3day & date <= lapse_onset) |>
  glimpse()

pred_1week <- labels |> 
  mutate(pred_1week = lapse_onset - days(7)) |> 
  inner_join(data, by = "subid") |>
  filter(date >= pred_1week & date <= lapse_onset) |>
  glimpse()
```

## EDA

### EDA on Contact

frequent contact based on self-reports
```{r}
data |> 
  ggplot(aes(x = freq_contacts)) +
  geom_bar() +
  facet_wrap(~subid)
```

```{r}
data |> 
  ggplot(aes(x = address)) +
  geom_bar() +
  facet_wrap(~subid)
```

## EDA on predictor windows

on all messages
```{r}
pred_1day |> 
  group_by(subid, lapse_onset) |> 
  summarize(n = n()) |>
  print_kbl()

pred_3day |> 
  group_by(subid, lapse_onset) |> 
  summarize(n = n()) |> 
  print_kbl()

pred_1week |> 
  group_by(subid, lapse_onset) |> 
  summarize(n = n()) |> 
  print_kbl()
```

for lapse label = yes only
```{r}
pred_1day |> 
  filter(label == "yes") |> 
  group_by(subid, lapse_onset) |> 
  summarize(n = n()) |> 
  print_kbl()

pred_3day |> 
  filter(label == "yes") |> 
  group_by(subid, lapse_onset) |> 
  summarize(n = n()) |> 
  print_kbl()

pred_1week |> 
  filter(label == "yes") |> 
  group_by(subid, lapse_onset) |> 
  summarize(n = n()) |> view()
  print_kbl()
```

on all messages by message type
```{r}
pred_1day |> 
  group_by(subid, lapse_onset, type) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = type, values_from = n) |> 
  print_kbl()

pred_3day |> 
  group_by(subid, lapse_onset, type) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = type, values_from = n) |> 
  print_kbl()

pred_1week |> 
  group_by(subid, lapse_onset, type) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = type, values_from = n) |> 
  print_kbl()
```

for lapse label = yes only by message type
```{r}
pred_1day |> 
  filter(label == "yes") |> 
  group_by(subid, lapse_onset, type) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = type, values_from = n) |> 
  print_kbl()

pred_3day |> 
  filter(label == "yes") |> 
  group_by(subid, lapse_onset, type) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = type, values_from = n) |> 
  print_kbl()

pred_1week |> 
  filter(label == "yes") |> 
  group_by(subid, lapse_onset, type) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = type, values_from = n) |> 
  print_kbl()
```

