---
title: "Clean up indiviudal messages"
author: "Coco Yu"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Code Status

## Notes

- Coco deleted subid 63, 64, 82, 100, 135, 190, 238, 211, 270, 116 (in mak_study_dates) because:
  - fewer than 100 messages

- Subid 1 & 84 were deleted (in mak_study_dates) because no important contact reported

- Coco filtered out only important contacts
  - excluding Self, Other, Co-Worker/Business Contact, Irrelevant/Spam 
  - including Parent, Friend, Sibling, Spouse/Significant Other, Child, Family-Other, Cousin, Grandparent, Counselor, Social Worker/Case Manager


## Setup

Chunk Defaults
```{r}
knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')
```

Conflicts
```{r}
options(conflicts.policy = "depends.ok")
```

```{r}
library(tidyverse)
library(kableExtra, exclude = "group_rows")
library(lubridate)
library(janitor, include.only = c("tabyl", "clean_names"))
library(here, include.only = c("here"))
```

Source Functions
```{r}
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
```

Absolute Paths 
```{r}
path_shared <- format_path(str_c("studydata/risk/data_processed/shared"))
path_messages <- format_path(str_c("studydata/risk/data_processed/messages"))
```

## Read in Data

```{r}
data <- read_csv(here(path_shared, "sms.csv"), 
                 col_types = cols(
                   subid = "d", start_study = "T", end_study = "T",
                   address = "c", date = "T", type = "f", body = "c",
                   msg_type = "f", phone_type = "f"
                 )) |> 
  mutate(date = as_datetime(date, tz = "America/Chicago"),
         start_study = as_datetime(start_study, tz = "America/Chicago"),
         end_study = as_datetime(end_study, tz = "America/Chicago"))
```

## Data Cleaning

filter out NA messages

```{r}
data |> filter(body == "￼") |> nrow()

data <- data |> 
  mutate(
    body = str_replace_all(body, "￼", ""),
    body = str_replace_all(body, "[\U0001F600-\U0001F64F\U0001F300-\U0001F5FF\U0001F680-\U0001F6FF\U0001F700-\U0001F77F\U0001F780-\U0001F7FF\U0001F800-\U0001F8FF\U0001F900-\U0001F9FF\U0001FA00-\U0001FA6F\U0001FA70-\U0001FAFF\U00002702-\U000027B0\U000024C2-\U0001F251]+", "")
  )

data <- data |> 
  filter(!is.na(body)) |> 
  filter(body != "")

subids <- data$subid |> 
  unique()
```

### Add Contact Information

#### Contact information based on self-reports

read in contact data
```{r}
contacts <- read_csv(here(path_shared, "contacts.csv"),
                     col_types = cols(.default = col_character(),
                                      subid = "d"))
```

check format of address in contacts.csv
```{r}
nchar(contacts$phone_number) |> 
  tabyl()

```

make address format consistent across contacts and data
```{r}
contacts <- contacts |> 
  mutate(
    phone_number = if_else(
      nchar(phone_number) == 10,
      paste0("1", phone_number),
      phone_number
    )
  ) |> 
  rename(address = phone_number) |> 
  select(subid, utc, address, contact_type)
```

resolve duplicated contacts

```{r}
contacts <- contacts |> 
  group_by(subid, address) |> 
  slice_max(utc, n = 1) |> 
  ungroup() |> 
  select(-utc)
```

see the list of contact type

```{r}
contacts |> 
  pull(contact_type) |> 
  unique()
```

combine data and contacts
```{r}
data <- contacts |> 
  right_join(data, by = c("subid", "address")) 
```


filter down important contact only

```{r}
data <- data |> 
  filter(
    str_detect(address, "~") |
    contact_type %in% c("Parent", "Friend", "Sibling", "Spouse/Significant Other", 
    "Child", "Family-Other", "Cousin", "Grandparent", "Counselor", 
    "Social Worker/Case Manager")
  )
```

```{r}
data <- data |> 
  select(-start_study, -end_study)
```

see what subids were excluded\
NOTE: The result should return NULL because subids (1, 84) were removed in mak_study_dates

```{r}
subids |> 
  setdiff(
    data |> 
      pull(subid) |> 
      unique()
  )
```


## EDA on full dataset

### EDA on number of messages

```{r}
data |> 
  tabyl(subid) |> 
  arrange(n) |> 
  print_kbl()
```

```{r}
data |> 
  group_by(subid) |> 
  summarise(n = n()) |>
  summarise(mean = mean(n),
            max = max(n),
            min = min(n),
            med = median(n))
```

### EDA on message length

get descriptives grouped by subid & pred_onset

```{r}
data |> 
  mutate(text_length = lengths(strsplit(data$body, "\\s+")))  |> 
  group_by(subid) |> 
  summarize(
    mean_length = mean(text_length),
    median_length = median(text_length),
    min_length = min(text_length),
    max_length = max(text_length)
  ) |> 
  print_kbl()
```

get means, median, min, max
```{r}

```

look into the long text

```{r}
# data |> 
#   mutate(text_length = lengths(strsplit(data$body, "\\s+")))  |> 
#   arrange(desc(text_length)) |> 
#   head(5) |> 
#   pull(body) |> 
#   print()
```

## Add lapse information

read in lapse labels
```{r}
labels <- read_csv(here(path_messages, "lapses.csv"), col_types = cols()) |> 
  mutate(day_start = as_datetime(day_start, tz = "America/Chicago"),
         day_end = as_datetime(day_end, tz = "America/Chicago"))
```

create prediction window
FIX: set relationship = many-to-many
```{r}

pred_1day <- labels |> 
  mutate(pred_onset = day_start - days(1)) |> 
  full_join(data, by = "subid") |>
  filter(date >= pred_onset & date <= day_start) 

pred_3day <- labels |> 
  mutate(pred_onset = day_start - days(3)) |> 
  full_join(data, by = "subid") |>
  filter(date >= pred_onset & date <= day_start) 

pred_1week <- labels |> 
  mutate(pred_onset = day_start - days(7)) |> 
  full_join(data, by = "subid") |>
  filter(date >= pred_onset & date <= day_start)
```

create dataframe for EDA: code text as NA if none message exists for one lapse label
```{r}
eda_1day <- pred_1day |> 
  full_join(
    labels |> 
      mutate(pred_onset = day_start - days(1)),
    by = c("subid", "day_start", "day_end", "pred_onset", "lapse")
  ) |> 
  mutate(body = if_else(is.na(body), "", body))

eda_3day <- pred_3day |> 
  full_join(
    labels |> 
      mutate(pred_onset = day_start - days(3)),
    by = c("subid", "day_start", "day_end", "pred_onset", "lapse")
  ) |> 
  mutate(body = if_else(is.na(body), "", body))

eda_1week <- pred_1week |> 
  full_join(
    labels |> 
      mutate(pred_onset = day_start - days(7)),
    by = c("subid", "day_start", "day_end", "pred_onset", "lapse")
  ) |> 
  mutate(body = if_else(is.na(body), "", body))
```

## EDA on predictor windows

histogram on number of messages (missing value included) per participant
```{r}
message_count_hist <- function(data){
  data |> 
    group_by(subid, pred_onset) |> 
    summarise(
      n = sum(body != ""),
      n_na = sum(body == "")
    ) |> 
    pivot_longer(
      cols = c(n, n_na), 
      names_to = "n", 
      values_to = "n_messages"
    ) |>
    ggplot(aes(x = pred_onset, y = n_messages, fill = n)) +
    geom_bar(stat = "identity") +
    facet_wrap(~subid, scales = "free_x") +
    theme_minimal()
}

message_count_hist(eda_1day)
message_count_hist(eda_3day)
message_count_hist(eda_1week)
```

**get basic summary information**\
missing: percentage of missingness (calculated by # of missing values divided by total number of lapses)\
total_messages: total count of messages\
mean/sd/min/max: descriptives of # of messages per label
```{r}
summary_info <- function(data){
  data |> 
    group_by(subid, pred_onset) |> 
    summarise(
      n = sum(body != ""),
      n_na = sum(body == "")
    ) |> 
    group_by(subid) |> 
    summarize(
      missing = sum(n_na) / n(),
      total_messages = sum(n) + sum(n_na),
      mean = mean(n),
      median = median(n),
      min = min(n),
      max = max(n)
    ) |> 
    pull(missing) |> 
    hist()
}

summary_info(eda_1day)
summary_info(eda_3day)
summary_info(eda_1week)
```

```{r}
eda_1week |> 
    group_by(subid, pred_onset) |> 
    summarise(
      n = sum(body != ""),
      n_na = sum(body == "")
    ) |> 
    group_by(subid) |> 
    summarize(
      missing = sum(n_na) / n(),
      total_messages = sum(n) + sum(n_na),
      mean = mean(n),
      median = median(n),
      min = min(n),
      max = max(n)
    ) |> 
  mutate(
    high_missing = if_else(missing > .2, 1, 0)
  ) |> 
  pull(high_missing) |> mean()
```

```{r}
missing_hist <- function(data){
  data |> 
    group_by(subid, pred_onset) |> 
    summarise(
      n_na = sum(body == "")
    ) |> 
    group_by(subid) |> 
    summarize(
      missing = sum(n_na),
      have_text = n() - sum(n_na)
    ) |> 
    pivot_longer(
      cols = c(missing, have_text), 
      names_to = "type", 
      values_to = "n"
    ) |> 
    mutate(subid = as.character(subid)) |> 
    ggplot(aes(y = subid, x = n, fill = type)) +
    geom_bar(stat = "identity") +
    theme_minimal()
}

missing_hist(eda_1day)
missing_hist(eda_3day)
missing_hist(eda_1week)
```

## Create Predictor Documents

function to combine prediction window
```{r}
combine_text <- function(df){
  df |> 
    select(subid, pred_onset, day_start, lapse, body) |> 
    group_by(subid, pred_onset) |> 
    summarize(body = ifelse(all(is.na(body)), NA_character_, 
                            str_c(body, collapse = " ")),
              lapse = first(lapse),
              .groups = "drop")
}
```

Combine all messages within same prediction window
```{r}
doc_1day <- combine_text(pred_1day)
doc_3day <- combine_text(pred_3day)
doc_1week <- combine_text(pred_1week)
```

Combine documents based no incoming/outgoing
```{r}
doc_incoming_1day <- combine_text(filter(pred_1day, type == "received"))
doc_incoming_3day <- combine_text(filter(pred_3day, type == "received"))
doc_incoming_1week <- combine_text(filter(pred_1week, type == "received"))

doc_outgoing_1day <- combine_text(filter(pred_1day, type == "sent"))
doc_outgoing_3day <- combine_text(filter(pred_3day, type == "sent"))
doc_outgoing_1week <- combine_text(filter(pred_1week, type == "sent"))
```

## Write-out csv

write out individual messages

```{r}
pred_1day |> 
  write_csv(here(path_messages, "predictors/pred_1day.csv"))
pred_3day |> 
  write_csv(here(path_messages, "predictors/pred_3day.csv"))
pred_1week |> 
  write_csv(here(path_messages, "predictors/pred_1week.csv"))

pred_1day |> 
  filter(type == "received") |> 
  write_csv(here(path_messages, "predictors/pred_incoming_1day.csv"))
pred_3day |> 
  filter(type == "received") |> 
  write_csv(here(path_messages, "predictors/pred_incoming_3day.csv"))
pred_1week |> 
  filter(type == "received") |> 
  write_csv(here(path_messages, "predictors/pred_incoming_1week.csv"))

pred_1day |> 
  filter(type == "sent") |> 
  write_csv(here(path_messages, "predictors/pred_outgoing_1day.csv"))
pred_3day |> 
  filter(type == "sent") |> 
  write_csv(here(path_messages, "predictors/pred_outgoing_3day.csv"))
pred_1week |> 
  filter(type == "sent") |> 
  write_csv(here(path_messages, "predictors/pred_outgoing_1week.csv"))
```

write out combined documents
```{r}
doc_1day <- doc_1day |> 
  write_csv(here(path_messages, "predictors/doc_1day.csv"))
doc_3day <- doc_3day |> 
  write_csv(here(path_messages, "predictors/doc_3day.csv"))
doc_1week <- doc_1week |> 
  write_csv(here(path_messages, "predictors/doc_1week.csv"))

doc_incoming_1day <- doc_incoming_1day |> 
  write_csv(here(path_messages, "predictors/doc_incoming_1day.csv"))
doc_incoming_3day <- doc_incoming_3day |> 
  write_csv(here(path_messages, "predictors/doc_incoming_3day.csv"))
doc_incoming_1week <- doc_incoming_1week |> 
  write_csv(here(path_messages, "predictors/doc_incoming_1week.csv"))

doc_outgoing_1day <- doc_outgoing_1day |> 
  write_csv(here(path_messages, "predictors/doc_outgoing_1day.csv"))
doc_outgoing_3day <- doc_outgoing_3day |> 
  write_csv(here(path_messages, "predictors/doc_outgoing_3day.csv"))
doc_outgoing_1week <- doc_outgoing_1week |> 
  write_csv(here(path_messages, "predictors/doc_outgoing_1week.csv"))
```

