---
title: "Make Lapse Labels"
author: "Coco Yu & Claire Punturieri"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Code Status

## Notes

- We dropped participants who didn't complete follow-up 1.

- John and Kendra have decided to drop subid 104's data for the following reasons:   
  - They had lapses every day on study except one day.    
  - Only had 75 surveys where a lapse was not reported.   
  - Viewing their lapse times it appears they were drinking constantly (morning and 
  night).   
  - They consistently report being uncertain that their goal is to be abstinent 
  (uncertain on 125 of 137 lapses. They also report they are uncertain in this goal 
  at followup 1 and 2.    
  - They are ultimately discontinued since they were struggling to gain sobriety.   
  - Unfortunately this drops 109 valid lapses.    


- John and Kendra have decided to drop subid 269's data for the following reasons:       
  - They completed 10-15 surveys on many days on study (avg number of surveys per 
  day is 6.76).  
  - Their responses indicate careless responding - they were filling 2-4 surveys out 
  within an hour of each other and the responses to the questions were completely different.     
  - They have questionable no lapse labels - they reported no lapses while on study but 
  according to notes left two messages for study staff where they admitted to drinking 
  over the weekend.   
  

- John and Kendra have decided to drop subid 204's data for the following reasons:    
  - Subid 204 had extremely poor compliance. 33 out of 89 study days had an EMA completed. 
  They only did a total of 5 surveys between followup 2 and 3.    
  - We don't trust their lapse labels - They report several lapses during their interviews 
  but times appear questionable (same time every night). They only report 1 lapse with EMA.
  - From notes - "Participant did not do many surveys during their second month of participation. 
  At their Follow-up 2 visit they reported several lapses that were not documented in their 
  EMAs - estimated lapse days/times in subid's raw data log."  
  - JC note: "There are issues with 204. They are missing lapses reported by interview. But they  
  also stopped doing any ema by 5/17 even though their study end date was 6/13. Probably need to 
  drop them for lapse analyses for anything after 5/17.  Probably also need to add in their 
  reported lapses at follow-up 2. OR we drop them at the end of follow-up 1 or wherever their 
  ema gets sketchy"
  
- hand-edit lapse_start for subid 213

## Setup

Chunk Defaults
```{r}
knitr::opts_chunk$set(attr.output='style="max-height: 500px;"')
```

Conflicts
```{r}
options(conflicts.policy = "depends.ok")
```

```{r}
library(tidyverse)
library(kableExtra, exclude = "group_rows")
library(lubridate)
library(janitor, include.only = c("tabyl", "clean_names"))
library(here, include.only = c("here"))
```

Source Functions
```{r}
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
```

Absolute Paths 
```{r}
path_shared <- format_path(str_c("studydata/risk/data_processed/shared"))
path_messages <- format_path(str_c("studydata/risk/data_processed/messages"))
```

## Read in Data

need further cleaning on lapses: for now, NA if NA in start_time
```{r}
visit_dates <- read_csv(here(path_messages, "study_dates.csv"),
                             col_types = cols()) |> 
  mutate(start_study = as_datetime(start_study, tz = "America/Chicago"),
         end_study = as_datetime(end_study, tz = "America/Chicago")) |> 
  glimpse()

lapses <- read_csv(here(path_shared, "lapses_day.csv"),
                   col_types = cols(subid = "d", exclude = "c",
                                    lapse_start_date = "c", 
                                    lapse_start_time = "c",
                                    lapse_end_date = "c",
                                    lapse_end_time = "c")) |> 
  mutate(lapse_start = as_datetime(lapse_start, tz = "America/Chicago")) |> 
  glimpse()

```


```{r}
labels <- visit_dates |> 
  rowwise() |>
  mutate(day_start = list(seq.POSIXt(start_study + days(1), end_study - days(1), by = "day"))) |> 
  unnest(day_start) |>
  mutate(day_start = update(day_start, hour = 4, minute = 0)) |> 
  select(subid, day_start) |> 
  glimpse()
```

clean up lapse start
```{r}
lapses <- lapses |> 
  mutate(
    lapse_end_date = if_else(is.na(lapse_end_date),
                             format(mdy(lapse_start_date) + days(1), "%m/%d/%Y"),
                             lapse_end_date),
    lapse_start_time = if_else(is.na(lapse_start_time), "20:00", lapse_start_time),
    lapse_end_time = if_else(is.na(lapse_end_time), "2:00", lapse_end_time),
    lapse_start = mdy_hm(paste(lapse_start_date, lapse_start_time)),
    lapse_start = force_tz(lapse_start, tzone = ema_1_6),
    lapse_start = as_datetime(lapse_start, tz = "America/Chicago"),
    lapse_end = mdy_hm(paste(lapse_end_date, lapse_end_time)),
    lapse_end = force_tz(lapse_end, tzone = ema_1_6),
    lapse_end = as_datetime(lapse_end, tz = "America/Chicago")
    ) |> 
  glimpse()

any(is.na(lapses$lapse_start))
any(is.na(lapses$lapse_end))
```

join lapses to labels\

```{r}
labels <- labels |> 
  mutate(day_end = day_start + days(1) - seconds(1),
         lapse = NA)

lapses_valid <- lapses |> 
  filter(exclude == "FALSE") |> 
  filter(!is.na(lapse_start))

for (i in 1:nrow(lapses_valid)){
  labels <- labels |> 
    mutate(lapse = if_else(subid == lapses_valid$subid[i] &
                             day_start <= lapses_valid$lapse_start[i] &
                             day_end >= lapses_valid$lapse_start[i],
                           "lapse", lapse))
}

lapses_exclude <- lapses |> 
  filter(exclude == "TRUE") |> 
  filter(!is.na(lapse_start))

for (i in 1:nrow(lapses_exclude)){
  labels <- labels |> 
    mutate(lapse = if_else(subid == lapses_exclude$subid[i] &
                             day_start <= lapses_exclude$lapse_start[i] &
                             day_end >= lapses_exclude$lapse_start[i],
                           "exclude", lapse),
           lapse = if_else(subid == lapses_exclude$subid[i] &
                             day_start <= lapses_exclude$lapse_end[i] &
                             day_end >= lapses_exclude$lapse_end[i],
                           "exclude", lapse))
}

labels <- labels |> 
  mutate(lapse = if_else(is.na(lapse), "no lapse", lapse)) |> 
  filter(lapse != "exclude")
```

## EDA
```{r}
(length(unique(labels$subid)))
```

```{r}
labels |> 
  group_by(subid, lapse) |> 
  summarize(n = n()) |> 
  print_kbl()
```

number of days by participant
```{r}
labels |> 
  group_by(subid) |>
  summarize(n_days = n()) |>
  ggplot(aes(x = subid, y = n_days)) +
  geom_point(size = 1) +
  geom_text(aes(label = subid), vjust = -0.5, hjust = 0.5) +
  theme_minimal()
```

number of lapses by participant
```{r}
labels |> 
  group_by(subid, lapse) |> 
  summarize(n_lapses = n()) |>
  filter(lapse == "lapse") |> 
  ggplot(aes(x = subid, y = n_lapses)) +
  geom_point(size = 0.5) +
  geom_text(aes(label = subid), vjust = -1, hjust = 0.5, size = 2) +
  theme_minimal()
```

percentage of lapses by participant
```{r}
labels |> 
  group_by(subid, lapse) |> 
  summarize(n_lapses = n()) |>
  full_join(labels |> group_by(subid) |> summarize(n_days = n())) |> 
  filter(lapse == "lapse") |> 
  ggplot(aes(x = subid, y = n_lapses/n_days)) +
  geom_point(size = 0.8) +
  geom_text(aes(label = subid), vjust = -1, hjust = 0.5, size = 2.5) +
  theme_minimal()
```

## Write-out csv

```{r}
labels |> 
  distinct() |> 
  write_csv(here(path_messages, "lapses.csv")) |> 
  glimpse()
```
